### 2025-10-28 factor-analysis.qmd
20250819 - updates
20250903 - updates to address updated data files
20251028 - additional info

Factor analysis involves the estimation of latent variables.
Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.
Factor analysis is a class of latent variable models that is designated to identify the structure of a measure or set of measures, and ideally, a construct or set of constructs.\index{factor analysis}
It aims to identify the optimal latent structure for a group of variables.\index{factor analysis}
The goal of factor analysis is to identify simple, parsimonious factors that underlie the "junk" (i.e., scores filled with measurement error) that we observe.\index{parsimony}\index{factor analysis!confirmatory}

@@

An additional consideration is the scaling of the variables: whether to use raw variables, standardized variables, or dividing some variables by a constant to make the variables' variances more similar.
Before performing a [principal component analysis](#sec-pca) (PCA), it is generally important to ensure that the variables included in the PCA are on the same scale.
PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
A common way of ensuring that variables are on the same scale is to standardize them using, for example, [*z* scores](#sec-zScores) that have a mean of zero and standard deviation of one.
By contrast, factor analysis can better accommodate variables that are on different scales.

@@

>>>>

Factor analysis involves the estimation of latent variables.
Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.
Factor analysis is a class of latent variable models that is designated to identify the structure of a measure or set of measures, and ideally, a construct or set of constructs.\index{factor analysis}
It aims to identify the optimal latent structure for a group of variables.\index{factor analysis}
The goal of factor analysis is to identify simple, parsimonious factors that underlie the "junk" (i.e., scores filled with measurement error) that we observe.\index{parsimony}\index{factor analysis!confirmatory}
That is, similar to [principal component analysis](#sec-pca), factor analysis can help us perform data reduction—i.e., it can help us reduce down a larger set of variables into a smaller set of factors.

@@

An additional consideration is the scaling of the variables: whether to use raw variables, standardized variables, or dividing some variables by a constant to make the variables' variances more similar.
Before performing a [principal component analysis](#sec-pca) (PCA), it is generally important to ensure that the variables included in the [PCA](#sec-pca) are on the same scale.
[PCA](#sec-pca) seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
A common way of ensuring that variables are on the same scale is to standardize them using, for example, [*z* scores](#sec-zScores) that have a mean of zero and standard deviation of one.
By contrast, factor analysis can better accommodate variables that are on different scales.

@@

lavaangui::plot_lavaan(efa_fit$nf3)
```
